---
title: "Assessing Vegetation Impacts from Deer Citizen Science Project: 2019 results"
author: "Matt Russell"
date: "`r Sys.Date()`"
output:
  pagedown::html_paged:
    toc: false
   #theme: lumen
    self_contained: false
    number_sections: true
---
# Overview
The AVID program is a citizen science programs that monitors the impacts of deer in Minnesota's forests. Citizen science volunteers collect important data that will allow researchers to gain insights into forest health.

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
setwd("C:/Users/russellm/Documents/Projects/AVID/AVID_reference")

seed<-read.csv("AVID_seedlings.csv")
plot<-read.csv("AVID_plots.csv")
site<-read.csv("AVID_sites.csv")

library(tidyverse)
library(maps)
library(mapdata)
library(ggrepel)
```

## AVID volunteers have measured many difference species that deer prefer to browse.
Here are the 15 species and their measured heights.

```{r, echo = F}
p.2018 <- ggplot(seed, aes(x=factor(Species), y=height1,fill=Species)) + 
  geom_boxplot()+xlab(" ")+ylab("Seedling height (in)")+
  theme(panel.background = element_rect(fill = "NA"),
        axis.line=element_line(color="black"),
        axis.title.x = element_text(size=12,color="black"),
        axis.title.y = element_text(size=12,color="black"),
        axis.text.x = element_blank(),
        axis.text.y  = element_text(size=12,color="black"))
p.2018
```


```{r, include=F}
# Create a lookup table for all species and merge to data:
  
  Species<-c("Balsam fir", "Black ash","Bur oak", "Green ash", "Ironwood", "Northern red oak", "Oak","Quaking aspen","Red maple","Red pine", "Sugar maple", "Swamp white oak", "White oak","White pine", "Yellow birch")
spp.group<-c("Pine/fir","Ash", "Oak", "Ash", "Other hardwoods", "Oak","Oak","Aspen","Maple","Pine/fir", "Maple", "Oak", "Oak","Pine/fir","Birch")
species.names<-data.frame(Species,spp.group)
  
seed<-merge(seed,species.names,by=c("Species"))
```

# Data from 2019
## We'll make a scatter plot that shows the measured heights.
Note we have grouped the species into broader species groups.

```{r, echo=F}
p.2018 <- ggplot(seed, aes(x=factor(spp.group), y=height1)) + 
  geom_point()+ geom_jitter()+
  xlab(" ")+ylab("Seedling height (inches)")+
  theme(panel.background = element_rect(fill = "NA"),
        axis.line=element_line(color="black"),
        axis.title.x = element_text(size=12,color="black"),
        axis.title.y = element_text(size=12,color="black"),
        axis.text.x = element_text(size=12,color="black",angle=45,hjust=1),
        axis.text.y  = element_text(size=12,color="black"))
p.2018
```

## Can we look at a specific volunteer's measurements?
Make a graph that shows one volunteers measurements relative to all of the data.

```{r, echo=F}
seed$Recorder_Name<-ifelse(seed$SiteID=="DD-NS","Your data!","Other AVID data")

p.2018 <- ggplot(seed, aes(x=factor(spp.group), y=height1,col=Recorder_Name)) + 
  geom_point()+ geom_jitter()+
  xlab(" ")+ylab("Seedling height (inches)")+
  theme(panel.background = element_rect(fill = "NA"),
        legend.title = element_blank(),
        legend.position = "top",
        legend.key=element_blank(),
        axis.line=element_line(color="black"),
        axis.title.x = element_text(size=12,color="black"),
        axis.title.y = element_text(size=12,color="black"),
       axis.text.x = element_text(size=12,color="black",angle=45,hjust=1),
        axis.text.y  = element_text(size=12,color="black"),
       legend.text = element_text(size=16,color="black"))
p.2018
```


# Which are the most common tree species measured?
Rank them by number of measurements.

```{r, echo=F}
avid.spp <-
  seed %>%
  group_by(Species) %>%
  summarize(n.seedlings=n(),
            mean.ht=round(mean(height1), 1),
            sd.ht=round(sd(height1), 1)) %>%
  arrange(desc(n.seedlings))

knitr::kable((avid.spp[,]), caption = 'The most common tree species measured.')
```

# Program impacts
## Make a map of where the data have been collected.
Data have been collected from 18 locations.

```{r, include=F}
# Map

states <- map_data("state")
mn_df <- subset(states, region == "minnesota")
counties <- map_data("county")
mn_county <- subset(counties, region == "minnesota")

mn_loc<-read.csv('C:\\Users\\russellm\\Documents\\Projects\\AVID\\AVID_reference\\mn_locations.csv')

all_base <- ggplot(data = mn_df, mapping = aes(x = long, y = lat, group=group)) + 
  coord_fixed(1.3) + 
  geom_polygon(color = "grey", fill = "white")+
  geom_polygon(color = "grey", fill = NA)+
  geom_polygon(data = mn_county, fill = NA, color = "grey") +
  geom_label_repel(data = mn_loc,aes(LON, LAT, label = Species),
                   box.padding = 0.1, point.padding = 0.1,
                   direction = "x",
                   nudge_x = 6, nudge_y = 2, angle=200,
                   fill = "gold", color = 'black', size = 2, inherit.aes=F) +
  xlab("") + ylab("")+
  theme(panel.background = element_rect(fill = "NA"),
        axis.line=element_blank(),
        axis.ticks=element_blank(),
        axis.title.x = element_text(size=12,color="black"),
        axis.title.y = element_text(size=12,color="black"),
        axis.text = element_blank())
```

```{r, echo=F}
all_base
```
